Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-03-28 Chapter-7/7-BackgroundSub.tex
s

So, how do we handle this contamination?

The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  
We estimate the percentage of electrons and muons in the beam via the beamline MC\footnote{Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
\begin{equation}
\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
\end{equation}
\begin{equation}
\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
\end{equation}
where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data. }.
Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...). The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.
The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.
We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} extremely small.

>>>>

So, how do we handle this contamination?
The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.

\subsection{Electron and Muon contamination}
We estimate the percentage of electrons and muons in the beam via the beamline MC. 
Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A.

\begin{figure}
\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam60A.png}
\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Chapter-7/Images/Beam100A.png}
\label{fig:BeamComposition}
\caption{Beam composition for the -60A runs (left) and -100A runs (right). The solid blue plot represents the pion content, the yellow plot represents the muon content and the grey plot represents the electron content. The plots are area normalized to the number of data events, shown in red. }
\end{figure}

Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4simulation, 
\begin{table}[]
\centering
\label{tab:beamline}
\begin{tabular}{|l|c|c|}
\hline
                     & I = -60 A           & I = -100 A \\ \hline
G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
\end{tabular}
\caption{Beamline composition per magnet settings}
\end{table}

\begin{table}[]
\centering
\label{tab:databreakdown}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
                                                     & I = -60 A          & I = -100 A   & Total     &  w$_{60A}$ &  w$_{100A}$\\ \hline
Data events after Mass Cut         &     70192          &  76056       & 146248 & 0.52 & 0.48\\ \hline
Data events for Cross Section     &                         &                   &              & & \\ \hline
\end{tabular}
\caption{Data events per magnet settings}
\end{table}

We calculate the electron to pion and muon to pion ratio on the whole sample as the weighted sum of the corresponding ratio in the two current settings, 
\begin{equation}
\frac{N_e}{N_\pi}_{Data} = w_{60A}\frac{N_e}{N_\pi}_{60A}  + w_{100A}\frac{N_e}{N_\pi}_{100A},
\end{equation}
\begin{equation}
\frac{N_\mu}{N_\pi}_{Data} = w_{60A}\frac{N_\mu}{N_\pi}_{60A}  + w_{100A}\frac{N_\mu}{N_\pi}_{100A},
\end{equation}
where the weights $w_{60A}$ and $w_{100A}$ are the percentage of events in the corresponding magnet configuration passing the mass selection in data, as shown in \ref{tab:databreakdown}.

Once the beam composition is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter, etc...). The percentage of electrons and muons surviving the selection chain is the  electron and muon contamination in the pion cross section sample.

\subsection{Contamination from secondaries}
The percentage of secondaries is given in the MC by the number of matched WC2TPC tracks which are not flagged as primary by Geant4.
We estimate the last type of contamination, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} extremely small.

### 2018-03-28 Chapter-8/8-TrackingOptimization.tex
s

>>>>

\section{Tracking spatial and angular resolution}
Scope of this study is understanding and comparing the tracking spatial and angular resolution on data and MC.
We start by selecting all the WC2TPC matched tracks. 
We fit a line on all the space points of the track and calculate the $\chi^2$. The $\chi^2$ distribution for data and MC is shown in Figure \ref{fig:Chi2AllPts}.

For the spatial and angular resolution study, we reject tracks with less than 14 space points. For each track, we order the space points according to their Z position and we split them in two sets: the first set counts all the points belonging to the first half of the track and the second set counts all the points belonging to the second half of the track. We remove the last 5 points in the first set and the first 5 points in the second set, so to have a gap in the middle of the original track. We fit the first and the second set of points with a line separately. We reject the event entirely if the  $\chi^2$ for the fit of either of the halves is greater than four.  We define a track middle plane as the plane perpendicular to the original track fit, positioned in the middle of its length. We project the tracks on the middle plane and calculate the impact parameter, $d$, i.e. the distance between the projected points. We also calculate the angle between the original track direction and the fit of the first and second half, called $\alpha_1$ and $\alpha_2$ respectively. The spatial resolution of the track will be $\sigma_S = \frac{d}{\sqrt 2}$ while the angular resolution of the tracks will be  $\sigma_\alpha = \alpha_1 - \alpha_2$. The distributions for data and MC for $\sigma_\alpha$ and $\sigma_S$ are given in \ref{fig:trackingResolution}.
